%%% Temporary changelog for supervisor review. This page is only printed
%%% while \showchangestrue is set in macros.tex; it is not part of the
%%% final thesis.

\chapter*{Changelog (revision of 15 July 2026)}

This revision responds to the fourth round of annotations. Passages added
or rewritten in this round are printed \changedD{in this green colour};
everything in black is unchanged from the previously annotated version
(the highlighting of the earlier rounds has been reverted to normal
text). Changes that cannot be highlighted in place, deletions and changes
outside the text, are listed here.

\subsection*{Chapter 5}
\begin{itemize}
  \item Section~5.3.2: the term \emph{intron} is no longer introduced.
        The duplicated $\max(\texttt{LS}, \texttt{LS})$ of Figure~5.10's
        activity tree is now described plainly as simplifying to
        \texttt{LS}, both in the text and in the figure caption.
  \item Section~5.3.3: ``the instances it failed on are not folded into
        the mean'' now reads ``are not included in the mean''; the second
        use of the same phrasing (``cannot be folded into the same
        six-cell ranking'') was aligned with it.
  \item Section~5.3.3: in the discussion of Figure~5.15, the
        parenthetical ``(an \emph{intron})'' was deleted; the code is
        described only as functionally neutral. The two remaining uses of
        ``load-bearing'' were reworded (``the two remaining references'',
        ``at important positions in the rule'').
  \item Section~5.4.2: the ten-class versus 108-class ARD\% shift is now
        called a \emph{sampling} effect rather than a composition effect.
  \item Section~5.4.2: a new paragraph after Table~5.29 adds the
        annotated observation on its feasibility rates: as RS grows, the
        NR-terminal condition's feasibility rises (on MMLIB100 from
        $81\%$ to $96\%$) while \texttt{baseline\_nr}'s does not progress,
        which is what the terminals' mechanism predicts once the
        non-renewable budget becomes the binding constraint. On MMLIB50
        both conditions rise at a similar rate, and the text says so.
\end{itemize}

\subsection*{Conclusion}
\begin{itemize}
  \item The clause ``which is what makes it the most instructive result
        of the three'' was deleted from the local-search verdict; the
        sentence now goes directly from the scale-specificity to the
        MMLIB100 outcome.
\end{itemize}

\subsection*{Attachments}
\begin{itemize}
  \item Appendix~A.1 now states what the electronic attachment (the ZIP
        archive uploaded to SIS) contains: the thesis PDF and the
        complete source code of the project, identical in content to the
        GitHub repository.
\end{itemize}
